---
title: "Adding or Modifying Parts of a Table"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gt)
library(tidyverse)
```


A **gt** table can contain a few useful parts for conveying additional information. These include a header (with a titles and subtitle), a footer (with footnotes and source notes), and additional areas for labels (row group labels, column spanner labels, the stubhead label). We can modify the look of table parts more generally with `tab_options()` and perform styling on targeted table locations with `tab_style()`.

### `tab_spanner()`: Add a spanner column label

``` r
tab_spanner(
  data,
  label,
  columns,
  gather = TRUE
)
```

Set a spanner column label by mapping it to columns already in the table. This label is placed above one or more column labels, spanning the width of those columns and column labels.

With `columns` we can use column names in double quotes (`"<column>"`), in `c()` (`c(<column>)`), or, we can use the following **tidyselect** expressions:

- `contains()`: contains a literal string
- `matches()`: matches a regular expression
- `starts_with()`: starts with a prefix
- `ends_with()`: ends with a suffix
- `everything()`: selects all columns

##### EXAMPLE

Let's use the `gtcars` table, but cut it down to size first:

```{r}
gtcars_small <- 
  gtcars %>%
  dplyr::select(
    -mfr, -trim, bdy_style, drivetrain,
    -drivetrain, -trsmn, -ctry_origin
  ) %>%
  dplyr::slice(1:8)

gtcars_small
```

------

Let's look at the table in **gt** so that we have a point of comparison.

```{r}
gtcars_small %>% gt(rowname_col = "model")
```

------

Use `gtcars` to create a **gt** table; Group several columns related to car performance under a spanner column with the label `performance`.

```{r}
gtcars_small %>%
  gt(rowname_col = "model") %>%
  tab_spanner(
    label = "performance",
    columns = c(hp, hp_rpm, trq, trq_rpm, mpg_c, mpg_h)
  )
```

------

### `tab_header()`: Add a table header

``` r
tab_header(
  data,
  title,
  subtitle = NULL
)
```

We can add a table header to the gt table with a title and even a subtitle. A table header is an optional table part that is positioned above the column labels. We have the flexibility to use Markdown or HTML formatting for the header's title and subtitle (with `md()` or `html()`).

##### EXAMPLE

Use `gtcars` to create a **gt** table; add a header part to contain a `title` and `subtitle`.

```{r}
gtcars %>%
  dplyr::select(mfr, model, msrp) %>%
  dplyr::slice(1:5) %>%
  gt() %>%
  tab_header(
    title = md("Data listing from **gtcars**"),
    subtitle = md("`gtcars` is an R dataset")
  )
```

------

### `tab_source_note()`: Add a source note citation

``` r
tab_source_note(
  data,
  source_note
)
```

We can add a source note to the footer part of any **gt** table. A source note is useful for citing the data included in the table. Several can be added, simply use multiple calls of `tab_source_note()` and they will be inserted in the order provided. We can use Markdown formatting for the note, or, if the table is intended for HTML output, we can include HTML formatting.

##### EXAMPLE

Use `exibble` to create a **gt** table. Add a source note to the table footer that cites the data source.

```{r}
exibble %>%
  gt() %>%
  tab_source_note(
    source_note = md("The `exibble` dataset is available in the **gt** package.")
  )
```

------

### `tab_footnote()`: Add a table footnote

``` r
tab_footnote(
  data,
  footnote,
  locations
)
```

The `tab_footnote()` function can make it a painless process to add a footnote to a **gt** table. There are two components to a footnote: (1) a footnote mark that is attached to the targeted cell text, and (2) the footnote text (that starts with the corresponding footnote mark) that is placed in the table's footer area. 

Each call of `tab_footnote()` will add a different note, and one or more cells can be targeted via the location helper (use in `locations`):

- `cells_title()` - target the table title or subtitle
- `cells_stubhead()` - target the table stubhead cell
- `cells_column_spanners()` - target the column spanners
- `cells_column_labels()` - target the column labels
- `cells_row_groups()` - target row groups
- `cells_stub()` - target cells in the table stub
- `cells_body()` - target data cells in the table body
- `cells_summary()` - target group summary cells
- `cells_grand_summary()` - target cells in a grand summary

Additionally, we can enclose several `cells_*()` calls within a `list()` if we wish to link the footnote text to different types of locations (e.g., body cells, row group labels, the table title, etc.).

##### EXAMPLE

Use `exibble` to create a **gt** table and then add a footnote to the `fctr` column label explaining what the short form means (`fctr` = 'factor').

```{r}
exibble %>%
  gt() %>%
  tab_footnote(
    footnote = "This is a factor column.",
    locations = cells_column_labels(columns = fctr)
  )
```

------

### `tab_style()`: Add custom styles to one or more cells

``` r
tab_style(
  data,
  style,
  locations
)
```

With the `tab_style()` function we can target specific cells and apply styles to them. 

This is done with the help of the following functions:

- `cell_text()`
- `cell_fill()`
- `cell_borders()`

For `locations` we use the `cells_*()` functions, just like in the `tab_footnote()` function. In the example below, we'll take things a step further with the `cells_body()` function and use a conditional statement in `rows` to target cells based on data.

##### EXAMPLES

Use a font from the Google Fonts service by using the `google_font()` function. Recommendations on some Google fonts can be found by using `info_google_fonts()`.

```{r}
exibble %>%
  dplyr::select(num, currency) %>%
  gt() %>%
  fmt_currency(columns = currency, currency = "EUR") %>%
  tab_style(
    style = cell_text(font = google_font("IBM Plex Sans"), weight = 500),
    locations = cells_body()
  )
```

------

Use `sp500` to create a **gt** table. Color entire rows of cells based on values in a particular column.

```{r}
sp500 %>%
  dplyr::filter(
    date >= "2015-12-01" &
    date <= "2015-12-15"
  ) %>%
  dplyr::select(-c(adj_close, volume)) %>%
  gt() %>%
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = cells_body(rows = close > open)
  ) %>%
  tab_style(
    style = list(
      cell_fill(color = "tomato"),
      cell_text(color = "white")
      ),
    locations = cells_body(rows = open > close)
  )
```

------

### `tab_options()`: Modify the table output options

``` r
tab_options(
  data,
  ...  <- huge argument list with dozens of options
)
```

We can modify the options available in a table. This function has a really huge set of arguments (but they are all documented in `?tab_options`, which helps).

##### EXAMPLES

Use `exibble` to create a **gt** table with all the main parts added; we can use this going forward to demo some `tab_options()`.

```{r}
tab_1 <- 
  exibble %>%
  dplyr::select(
    -c(fctr, date, time, datetime)
  ) %>%
  gt(
    rowname_col = "row",
    groupname_col = "group"
  ) %>%
  tab_header(
    title = md("Data listing from **exibble**"),
    subtitle = md("`exibble` is an R dataset")
  ) %>%
  fmt_number(columns = num, locale = "de") %>%
  fmt_currency(columns = currency, currency = "EUR", locale = "de") %>%
  tab_footnote(
    footnote = md("Numbers in `grp_b`."),
    locations = cells_body(
      columns = num,
      rows = num > 1000
    )
  ) %>%
  tab_footnote(
    footnote = "Very high amounts.",
    locations = cells_body(
      columns = currency,
      rows = currency > 1000)
  ) %>%
  tab_footnote(
    footnote = "Alphabetical fruit.",
    locations = cells_column_labels(
      columns = char
    )
  )
```

------

Modify the table width (with `table.width`) to `100%` (which spans the entire content width area).

```{r}
tab_1 %>%
  tab_options(
    table.width = pct(100) # pct() helper function used here
  )
```

------

Modify the table's background color (with `table.background.color`) to be `"lightcyan"`.

```{r}
tab_1 %>%
  tab_options(
    table.background.color = "lightcyan" # R and hexadecimal colors can be used 
  )
```

------

Use letters as the glyphs for footnote references (with `footnotes.marks` and the `letters` vector).

```{r}
tab_1 %>%
  tab_options(
    footnotes.marks = letters
  )
```

------

Change the padding of data rows to `5px` with `data_row.padding`.

```{r}
tab_1 %>% 
  tab_options(
    data_row.padding = px(5) # the `px()` helper function is used here
  ) 
```

------

Reduce the size of the title and the subtitle text (with `heading.title.font.size` and `heading.subtitle.font.size`).

```{r}
tab_1 %>%
  tab_options(
    heading.title.font.size = "small",
    heading.subtitle.font.size = "x-small"
  )
```

------
