---
title: "Formatting Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gt)
library(tidyverse)
```


Columns of data can be formatted with the `fmt_*()` functions. We can specify the rows of these columns quite precisely with the `rows` argument. We get to apply these functions exactly once to each data cell (last call wins).

### `fmt_number()`: Format numeric values

``` r
fmt_number(
  data,
  columns,
  rows = everything(),
  decimals = 2,
  n_sigfig = NULL,
  drop_trailing_zeros = FALSE,
  drop_trailing_dec_mark = TRUE,
  use_seps = TRUE,
  scale_by = 1,
  suffixing = FALSE,
  pattern = "{x}",
  sep_mark = ",",
  dec_mark = ".",
  locale = NULL
)
```

With numeric values in a **gt** table, we can perform number-based formatting so that the targeted values are rendered with a higher consideration for tabular presentation.

##### EXAMPLES

Use `exibble` to create a **gt** table. Format the `num` column using `fmt_number()` and its default options.

```{r}
exibble %>%
  dplyr::select(num) %>%
  gt() %>%
  fmt_number(columns = num)
```

------

Again with `exibble`, format the `num` column as numeric with three decimal places. Don't use digit separators (`use_seps = FALSE`).

```{r}
exibble %>%
  dplyr::select(num) %>%
  gt() %>%
  fmt_number(
    columns = num,
    decimals = 3,
    use_seps = FALSE
  )
```

------

Format only the first three rows of `num`, this time with 4 decimal places.

```{r}
exibble %>%
  dplyr::select(num) %>%
  gt() %>%
  fmt_number(
    columns = num,
    rows = 1:3,
    decimals = 4
  )
```

------

Format only the rows of `num` (to 4 decimal places) where values in `currency` are greater than `30`.

```{r}
exibble %>%
  dplyr::select(num, currency) %>%
  dplyr::arrange(currency) %>%
  gt() %>%
  fmt_number(
    columns = num,
    rows = currency > 30,
    decimals = 4
  )
```

------

Use `countrypops` to create a **gt** table. Format all numeric columns to use large-number suffixing (with `suffixing = TRUE`).

```{r}
countrypops %>%
  dplyr::select(country_code_3, year, population) %>%
  dplyr::filter(country_code_3 %in% c("CHN", "IND", "USA", "PAK", "IDN")) %>%
  dplyr::filter(year > 1975 & year %% 5 == 0) %>%
  tidyr::spread(year, population) %>%
  dplyr::arrange(desc(`2015`)) %>%
  gt(rowname_col = "country_code_3") %>%
  fmt_number(
    columns = 2:9, # <- selecting columns 2 to 9 in *passed in dataset*
    decimals = 2,
    suffixing = TRUE,
    locale = "es" # <- try: "fr_FR", "de", "de_AT" (i.e., <lang>_<region>)
  )
```

------

### `fmt_currency()`: Format values as currencies

``` r
fmt_currency(
  data,
  columns,
  rows = everything(),
  currency = "USD",
  use_subunits = TRUE,
  accounting = FALSE,
  decimals = NULL,
  drop_trailing_dec_mark = TRUE,
  use_seps = TRUE,
  scale_by = 1,
  suffixing = FALSE,
  pattern = "{x}",
  sep_mark = ",",
  dec_mark = ".",
  placement = "left",
  incl_space = FALSE,
  locale = NULL
)
```

With numeric values in a **gt** table, we can perform currency-based formatting. This function supports both automatic formatting with a three-letter or numeric currency code. We can also specify a custom currency that is formatted according to the output context with the `currency()` helper function. Numeric formatting facilitated through the use of a locale ID.

##### EXAMPLES

Use `exibble` to create a **gt** table. Keep only the `num` and `currency`, columns, then, format those columns using the `CNY` and `GBP` currencies.

```{r}
exibble %>%
  dplyr::select(num, currency) %>%
  gt() %>%
  fmt_currency(
    columns = num,
    currency = "CNY"
  ) %>%
  fmt_currency(
    columns = currency,
    currency = "GBP"
  )
```

------

### `fmt_markdown()`: Format Markdown text

``` r
fmt_markdown(
  data,
  columns,
  rows = everything()
)
```

Any Markdown-formatted text in the incoming cells will be transformed to the appropriate output type during render when using `fmt_markdown()`.

##### EXAMPLE

Create a few Markdown-based text snippets.

```{r}
text_1a <- "
### This is Markdown.

Markdown’s syntax is comprised entirely of
punctuation characters, which punctuation
characters have been carefully chosen so as
to look like what they mean... assuming
you’ve ever used email.
"

text_1b <- "
Info on Markdown syntax can be found
[here](https://daringfireball.net/projects/markdown/).
"

text_2a <- "
The **gt** package has these datasets:

 - `countrypops`
 - `sza`
 - `gtcars`
 - `sp500`
 - `pizzaplace`
 - `exibble`
"

text_2b <- "
There's a quick reference [here](https://commonmark.org/help/).
"
```

Arrange the text snippets as a tibble using the `dplyr::tribble()` function. Then, create a **gt** table and format all columns with `fmt_markdown()`.

```{r}
dplyr::tribble(
  ~Markdown, ~md,
  text_1a,   text_2a,
  text_1b,   text_2b,
) %>%
  gt() %>%
  fmt_markdown(columns = everything()) %>%
  tab_options(table.width = px(400))
```

------

### `fmt_missing()`: Format missing values

``` r
fmt_missing(
  data,
  columns,
  rows = everything(),
  missing_text = "---"
)
```

Wherever there is missing data (i.e., `NA` values) a customized mark may present better than the standard `NA` text that would otherwise appear. The `fmt_missing()` function allows for this replacement through its `missing_text` argument (where an em dash serves as the default).

##### EXAMPLE

Use `exibble` to create a **gt** table. The `NA` values in different columns will be given replacement text.

```{r}
exibble %>%
  dplyr::select(-row, -group) %>%
  gt() %>%
  fmt_missing(
    columns = 1:2,
    missing_text = md("**missing**")
  ) %>%
  fmt_missing(
    columns = 4:7,
    missing_text = md("_**UDL**_")
  )
```

------

### `data_color()`: Set data cell colors using a palette or a color function

``` r
data_color(
  data,
  columns,
  colors,
  alpha = NULL,
  apply_to = c("fill", "text"),
  autocolor_text = TRUE
)
```

It's possible to add color to data cells according to their values. The `data_color()` function colors all rows of any `columns` supplied. There are two ways to define how cells are colored: (1) through the use of a supplied color palette, and (2) through use of a color mapping function available from the **scales** package.

The first method colorizes cell data according to whether values are character or numeric. The second method provides more control over how cells are colored since we provide an explicit color function and thus other requirements such as bin counts, cut points, or a numeric domain.

##### EXAMPLES

Use `countrypops` to create a **gt** table. Apply a color scale to the `population` column with `scales::col_numeric`, four supplied colors, and a domain.

```{r}
countrypops %>%
  dplyr::filter(country_name == "Mongolia") %>%
  dplyr::select(-contains("code")) %>%
  tail(10) %>%
  gt() %>%
  data_color(
    columns = population,
    colors = scales::col_numeric(
      palette = c("red", "orange", "green", "blue"),
      domain = c(2E6, 4E6)
    )
  )
```

------
