---
title: "Exploring the films dataset"
format: html
---

```{r}
#| include: false
library(pointblank)
library(tidyverse)
library(beepr)
```

# First look at the `films` dataset

We are going to load in my films dataset that I assembled from my Letterboxd
data (I have an account, my profile is https://letterboxd.com/rich_i/)

We will use `readr::read_csv()` for this and I set the `col_types` value using
what I know about the dataset

```{r}
films <-
  readr::read_csv(
    file = "2026-08-04-DSLab-pointblank-for-R/films.csv",
    col_types = cols(
      film_title = col_character(),
      year = col_integer(),
      letterboxd_uri = col_character(),
      film_rating = col_double(),
      director = col_character()
    )
  )
```

Look at the tibble.

```{r}
films
```

We're in Positron, so we can use `View()` to quickly see the dataset in the
Data Viewer.

```{r}
View(films)
```

Just from looking at the stats in the Data Viewer we see issues. We can also use
summary:

```{r}
summary(films)
```

Not as good though! Doesn't show missing values for character columns :(

One thing that pointblank offers is `scan_data()`. It can provide a lot of useful
information about a dataset (especially in its `Variables` view). Use this
to get an HTML-based report:

```{r}
pointblank::scan_data(films, sections = "OVS")
```

Some of what's in there is not useful, but some of it really is!

# Column descriptions AKA a data dictionary

Columns and what they mean (a data dictionary)

`film_title`
The name of the film. Punctuation and special characters are fine. This is
typically the English name of the film (or at least transliterated to English).
Every film should have a title.

`year`
The year the film was widely released. Every film has a single year value.

`letterboxd_uri`
A shortened link to the Letterboxd film page for `film_title`. This is publicly
accessible (no login required). Every film should have a value for this.

`film_rating`
This is what I rated the film in Letterboxd through my personal account at
https://letterboxd.com/rich_i/. Every film should be rated.

`director`
The name of the director usually as '{first_name} {last_name}' (but not always,
and it's fine). Sometimes there are multiple directors and each are separated
with `", "`. There should always be at least one director.

The problem: seems like some looking at the data shows weird values on occasion!

The solution: let's validate!

# Using pointblank

The workflow for data validation in **pointblank** involves three basic components:

- the creation of an 'agent' (this is the main data collection and reporting object)
- the declaration of validation steps using validation functions (as many as you need)
- the interrogation of the data (here the agent finally carries out the validation tasks)

The validation sandwich:

```r
agent <-
  create_agent(...) |>
  << validation functions >> |>
  interrogate()
```

Validation functions are documented in here:

https://rstudio.github.io/pointblank/reference/index.html#validation-expectation-and-test-functions

Okay let's now try to validate this dataset:

```{r}

```


# Splitting a Validated Dataset

After a validation, you might want to split the dataset rows to keep good parts,
and to look more closely at bad parts (failing test units). We use the function
called `get_sundered_data()` for this.

```{r}

```


# Using `draft_validation()` to get up and running quickly

You can use a function called `draft_validation()`, to get a large piece of
validation code.

```{r}

```
