---
title: "Creating a Table, Exporting"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gt)
library(tidyverse)
```


All **gt** tables begin with a call to `gt()`, where we would supply the input data table (data frame or tibble) and some basic options for creating a stub (`rowname_col`) and row groups (`groupname_col`). We can also introduce a grouped tibble (via **dplyr**’s `group_by()`) for more precise divisions of rows into row groups. 

The `gt_preview()` function is great for getting a nicely-formatted preview of a data table (defaulting to the first 5 rows and the last row).

Finally, sometimes we want to export the table to some format. It could be HTML, it could be RTF... whichever you go with, you can use the `gtsave()` function.

### `gt()`: Create a **gt** table object

``` r
gt(
  data,
  rowname_col = "rowname",
  groupname_col = dplyr::group_vars(data),
  rownames_to_stub = FALSE,
  auto_align = TRUE,
  id = NULL,
  row_group.sep = getOption("gt.row_group.sep", " - ")
)
```

The `gt()` function creates a **gt** table object when provided with table data. Using this function is the first step in a typical **gt** workflow. Once we have the **gt** table object, we can perform styling transformations before rendering to a display table of various formats.

------

##### EXAMPLES

The **gt** package contains a few datasets:

- `countrypops` - Yearly populations of countries from 1960 to 2017
- `sza` - Twice hourly solar zenith angles by month & latitude
- `gtcars` - Deluxe automobiles from the 2014-2017 period
- `sp500` - Daily S&P 500 Index data from 1950 to 2015
- `pizzaplace` - A year of pizza sales from a pizza place
- `exibble` - A toy example tibble for testing with **gt**: `exibble`

Let's print out `exibble` and see what it looks like:

```{r}
exibble
```

Now lets introduce `exibble` to the `gt()` function.

```{r}
exibble %>% gt()
```

------

We can modify the layout a bit in the first step and create a stub. To do this, we supply the name of the column we'd like to use as the stub to the `rowname_col` argument.

```{r}
exibble %>% gt(rowname_col = "row")
```

------

We can do some more structuring. Sets of rows can be grouped, showing a *row group label* above each group. We can do this by using a column containing categorical values (usually grouping labels). Let's use the `rowname_col` argument of `gt()` and see what row groups look like:

```{r}
exibble %>% gt(rowname_col = "row", groupname_col = "group")
```

------

Even if rows are in a weird order, **gt** will put the rows in the correct group (but still respect the order of rows).

```{r}
exibble %>% 
  dplyr::sample_n(size = 8) %>%
  gt(rowname_col = "row", groupname_col = "group")
```

------

Say you have a data frame with row names (like `mtcars`). You can put those row.names in the stub with `rownames_to_stub = TRUE`:

```{r}
mtcars %>%
  dplyr::slice_head(n = 10) %>%
  gt(rownames_to_stub = TRUE)
```

------

### `gt_preview()`: Preview a **gt** table object

``` r
gt_preview(
  data,
  top_n = 5,
  bottom_n = 1,
  incl_rownums = TRUE
)
```

Sometimes you may want to see just a small portion of your input data. We can use `gt_preview()` in place of `gt()` to get the first *x* rows of data and the last *y* rows of data (which can be set by the `top_n` and `bottom_n` arguments).

##### EXAMPLES

Use `gtcars` to create a **gt** table preview (with only a few of its columns). You'll see the first five rows and the last row.

```{r}
gtcars %>%
  dplyr::select(mfr, model, year) %>%
  gt_preview()
```

------

Don't want to see the row numbers? That's fine, use `incl_rownums = FALSE`.

```{r}
sp500 %>%
  gt_preview(
    bottom_n = 5,
    incl_rownums = FALSE
  )
```

------

### `gtsave()`: Save a **gt** table as a file

``` r
gtsave(
  data,
  filename,
  path = NULL,
  ...
)
```

The `gtsave()` function makes it easy to save a **gt** table to a file. The function assumes the output file type by the extension provided in the output filename. This will produce either an HTML, PDF, PNG, LaTeX, or RTF file.

##### EXAMPLES

Use `exibble` to create a **gt** table.

```{r}
gt_tbl <-
  exibble %>%
  gt(
    rowname_col = "row",
    groupname_col = "group"
  )
```

------

Get an HTML file with inlined CSS (which is necessary for including the table as part of an HTML email in the **blastula** package).

```{r}
# gt_tbl %>% gtsave("tab_1_inlined.html", inline_css = TRUE)
```

By leaving out the `inline_css` option, we get a more conventional HTML file with embedded CSS styles.

```{r}
# gt_tbl %>% gtsave("tab_1.html")
```

Get an RTF file by using the .rtf extension in the `filename`.

```{r}
# gt_tbl %>% gtsave("tab_r.rtf")
```

