---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gt)
library(readr)
```

## Data preparation work before generating the summary table

Import the data table.

```{r}
tbl <- read_rds(file = "data-14_3_13.rds")
```

Take a look at the structure of the table.

```{r}
tbl
```

Get the current date with the `Sys.Date()` function.

```{r}
reporting_date <- as.character(Sys.Date())
```

Get all rows with the *N* values.

```{r}
n_value_rows <- tbl %>% filter(label == "n")
```

Pull the highest *N* value for each of the required columns.

```{r}
placebo_n <- n_value_rows %>% pull(placebo) %>% max()
xanomeline_ld_n <- n_value_rows %>% pull(xanomeline_ld) %>% max()
xanomeline_hd_n <- n_value_rows %>% pull(xanomeline_hd) %>% max()
```

## Summary table generation with **gt**

This table is built up using many of the same functions we used in `gt-01-14_2_01.Rmd`.

```{r}
gt_table_final <-
  tbl %>%
  gt() %>%
  tab_header(
    title = "Table 14.3.13",
    subtitle = "CIBIC+ - Categorical Analysis - LOCF",
    preheader = c("Protocol: CDISCPILOT01", "Population: Efficacy")
  ) %>%
  sub_missing(columns = everything(), missing_text = "") %>%
  fmt_number(columns = p, decimals = 4) %>%
  fmt_percent(
    columns = ends_with("_pct"),
    decimals = 0,
    scale_values = FALSE
  ) %>%
  cols_merge_n_pct(col_n = placebo, col_pct = placebo_pct) %>%
  cols_merge_n_pct(col_n = xanomeline_ld, col_pct = xanomeline_ld_pct) %>%
  cols_merge_n_pct(col_n = xanomeline_hd, col_pct = xanomeline_hd_pct) %>%
  cols_label(
    category = "",
    label = "Assessment",
    placebo  = md(paste0("Placebo  \n(N=", placebo_n, ")")),
    xanomeline_ld = md(paste0("Xanomeline  \nLow Dose  \n(N=", xanomeline_ld_n, ")")),
    xanomeline_hd = md(paste0("Xanomeline  \nHigh Dose  \n(N=", xanomeline_hd_n, ")")),
    p = "p-value"
  ) %>%
  tab_footnote(
    footnote = "Overall comparison of treatments using CMH test (Pearson Chi-Square), controlling for site group.",
    locations = cells_column_labels(columns = p)
  ) %>%
  tab_source_note(
    source_note = reporting_date
  ) %>%
  cols_width(
    1 ~ pct(10),
    2 ~ pct(25)
  ) %>%
  tab_options(
    page.orientation = "landscape",
    page.numbering = TRUE,
    page.header.use_tbl_headings = TRUE,
    page.footer.use_tbl_notes = TRUE
  )

gt_table_final
```

Writing the table to HTML can be done with `gtsave()`.

```{r}
gt_table_final %>% gtsave("html_14.3.13.html")
```

Write the **gt** table to an RTF document.

```{r}
gt_table_final %>% gtsave("tbl_14.3.13.rtf")
```
