---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gt)
library(readr)
```

## Data preparation work before generating the summary table

Import the data table.

```{r}
tbl <- read_rds(file = "data-14_5_01.rds")
```

Take a look at the structure of the table.

```{r}
tbl
```

Get the current date with the `Sys.Date()` function.

```{r}
reporting_date <- as.character(Sys.Date())
```

Get all rows with the *N* values.

```{r}
n_value_rows <- tbl %>% filter(label == "n")
```

Pull the highest *N* value for each of the required columns.

```{r}
placebo_n <- 86L
xanomeline_ld_n <- xanomeline_hd_n <- 84L
```

## Summary table generation with **gt**

This table is built up using many of the same functions we used in `gt-01-14_2_01.Rmd`. There are some notable differences though:

- the `fmt_integer()` function is being used to format numeric values as integers, and, we are wrapping each value in square brackets with `pattern = "[{x}]"`
- there are two `fmt_numeric()` statements that format p-values and they use a `rows` expression (`fisher_placebo_v_xanomeline_ld < 0.2`) to target values less than `0.2` and the `pattern = "{x}*"` will apply the asterisk to the p-value in these cases
- there is another set of `fmt_numeric()` statements that replace values close to `1` with `>0.99` by using a method similar to the above (`rows` expression, and a `pattern`)
- the `tab_spanner()` function is used to create spanner column labels (these are labels over top column labels)

```{r}
tbl <-
  tbl %>%
  mutate(label = ifelse(!is.na(label), paste0("  ", label), label))

gt_table_final <-
  tbl %>%
  gt() %>%
  tab_header(
    title = "Table 14-5.01",
    subtitle = "Incidence of Treatment Emergent Adverse Events by Treatment Group",
    preheader = c("Protocol: CDISCPILOT01", "Population: Safety")
  ) %>%
  sub_missing(columns = everything(), missing_text = "") %>%
  fmt_number(
    columns = c(fisher_placebo_v_xanomeline_ld, fisher_placebo_v_xanomeline_hd),
    decimals = 4
  ) %>%
  fmt_percent(
    columns = ends_with("_pct"),
    decimals = 1,
    scale_values = FALSE
  ) %>%
  fmt_integer(
    columns = ends_with("_ae"),
    pattern = "[{x}]"
  ) %>%
  fmt_number(
    columns = starts_with("fisher"),
    decimals = 3
  ) %>%
  fmt_number(
    columns = fisher_placebo_v_xanomeline_ld,
    rows = fisher_placebo_v_xanomeline_ld < 0.2,
    decimals = 3,
    pattern = "{x}*"
  ) %>%
  fmt_number(
    columns = fisher_placebo_v_xanomeline_hd,
    rows = fisher_placebo_v_xanomeline_hd < 0.2,
    decimals = 3,
    pattern = "{x}*"
  ) %>%
  fmt_number(
    columns = fisher_placebo_v_xanomeline_ld,
    rows = fisher_placebo_v_xanomeline_ld > 0.99,
    pattern = ">.99"
  ) %>%
  fmt_number(
    columns = fisher_placebo_v_xanomeline_hd,
    rows = fisher_placebo_v_xanomeline_hd > 0.99,
    pattern = ">.99"
  ) %>%
  cols_merge_n_pct(col_n = placebo_n, col_pct = placebo_pct) %>%
  cols_merge_n_pct(col_n = xanomeline_ld_n, col_pct = xanomeline_ld_pct) %>%
  cols_merge_n_pct(col_n = xanomeline_hd_n, col_pct = xanomeline_hd_pct) %>%
  tab_spanner(
    label = md(paste0("Placebo  \n(N=", placebo_n, ")")),
    columns = starts_with("placebo")
  ) %>%
  tab_spanner(
    label = md(paste0("Xanomeline  \nLow Dose  \n(N=", xanomeline_ld_n, ")")),
    columns = starts_with("xanomeline_ld")
  ) %>%
  tab_spanner(
    label = md(paste0("Xanomeline  \nHigh Dose  \n(N=", xanomeline_hd_n, ")")),
    columns = starts_with("xanomeline_hd")
  ) %>%
  tab_spanner(
    label = md("Fisher's Exact  \np-values"),
    columns = starts_with("fisher")
  ) %>%
  cols_label(
    placebo_n = "n(%)",
    xanomeline_ld_n = "n(%)",
    xanomeline_hd_n = "n(%)",
    placebo_ae = "[AEs]",
    xanomeline_ld_ae = "[AEs]",
    xanomeline_hd_ae = "[AEs]",
    fisher_placebo_v_xanomeline_ld = md("Placebo  \nvs.  \nLow Dose"),
    fisher_placebo_v_xanomeline_hd = md("Placebo  \nvs.  \nHigh Dose")
  ) %>%
  tab_footnote(footnote = "Note: Treatment emergent events are defined as events which start on or after the start of treatment.") %>%
  tab_footnote(footnote = "Note: Adverse events are coded using MedDRA.") %>%
  tab_footnote(footnote = "Note: Percentages are based on the number of subjects in the safety population within each treatment group.") %>%
  tab_footnote(footnote = "Note: P-values are based on Fisher's Exact test for the comparison of placebo versus each active treatment group. An asterisk is appended to p-values that are less than 0.15.") %>%
  tab_footnote(footnote = "Note: The column [AE] represents the total number of times an event was recorded.") %>%
  tab_source_note(
    source_note = paste('Program Source: t-14-5-01.R     Executed: (Draft)', reporting_date)
  ) %>%
  cols_width(
    1 ~ pct(30),
    ends_with("n") ~ pct(34/3),
    ends_with("ae") ~ pct(18/3),
    starts_with("fisher") ~ pct(18/2)
  ) %>%
  tab_options(
    page.orientation = "landscape",
    page.numbering = TRUE,
    page.header.use_tbl_headings = TRUE,
    page.footer.use_tbl_notes = TRUE
  )

gt_table_final
```

Writing the table to HTML can be done with `gtsave()`.

```{r}
gt_table_final %>% gtsave("html_14.5.01.html")
```

Write the **gt** table to an RTF document.

```{r}
gt_table_final %>% gtsave("tbl_14.5.01.rtf")
```
