---
title: "Exibble Example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gt)
library(tidyverse)
```

Have a look at `exibble`.

```{r}
exibble
```

View an `exibble` **gt** table

```{r}
exibble %>% gt()
```

Create a stub (with row labels), and, include row groups (using the `row` and `group` columns).

```{r}
tab <-
  exibble %>%
  gt(
    rowname_col = "row",
    groupname_col = "group"
  )

tab
```

Format the `num` column with **gt**'s scientific notation.

```{r}
tab_2 <-
  tab %>%
  fmt_scientific(
    columns = vars(num),
    decimals = 3)

tab_2
```

View a table of date styles.

```{r}
info_date_style()
```

Format the dates in `date` (but only those rows that satisfy a condition) with `date_style` 6.

```{r}
tab_3 <-
  tab_2 %>%
  fmt_date(
    columns = vars(date),
    rows = grepl("^[a-d]", char),
    date_style = 6
  )

tab_3
```

Remove some columns from the table.

```{r}
tab_4 <-
  tab_3 %>%
  cols_hide(
    columns = vars(char, fctr, time, datetime))

tab_4
```

Format the currency column to have values in pounds.

```{r}
tab_5 <-
  tab_4 %>%
  fmt_currency(
    columns = vars(currency),
    currency = "GBP"
    )

tab_5
```

Add conditional footnotes to some of the `currency` values.

```{r}
tab_6 <-
  tab_5 %>%
  tab_footnote(
    footnote = "These are lower prices.",
    locations = cells_data(
      columns = vars(currency),
      rows = currency < 20
    )
  )

tab_6
```

Add a footnote to the `currency` column label itself.

```{r}
tab_7 <-
  tab_6 %>%
  tab_footnote(
    footnote = "All values are in GBP.",
    locations = cells_column_labels(
      columns = vars(currency)
    )
  )

tab_7
```

Add a header to the table (with a title and a subtitle).

```{r}
tab_8 <-
  tab_7 %>%
  tab_header(
    title = "An Exibble Example",
    subtitle = md("Uses the `exibble` dataset from **gt**")
  )

tab_8
```

Have the footnote glyphs be a sequence of lowercase letters.

```{r}
tab_9 <-
  tab_8 %>%
  tab_options(footnote.glyph = letters)

tab_9
```

Color the stub cells in blue.

```{r}
tab_10 <-
  tab_9 %>%
  tab_options(column_labels.background.color = "blue")

tab_10
```
